---
title: "GRACKLE Simulation Tutorial"
author: "GRACKLE Team"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This tutorial demonstrates how to simulate gene expression data and evaluate the GRACKLE model using a preprocessed breast network. The steps include generating random graphs, adjusting transitivity, adding noise, and performing GRACKLE grid search.

## Load Required Libraries

```{r load-libraries}
library(tidyverse)
library(igraph)
library(devtools)
library(parallel)
library(optparse)
load_all()
library(reticulate)
use_virtualenv("/mnt/grackle_env")
library(tensorflow)
```

## Load Preprocessed Breast Network

```{r load-network}
breast_graph <- get(load("../GRACKLE_data/data/breast_igraph_prob_1_cor_0_05.RData"))
```

## Define Parameters

```{r define-parameters}
noise_sequence <- seq(0.3, 0.7, 0.2)
transitivity_sequence <- c(0, seq(0.1, 0.3, 0.1))
num_nodes <- 400
num_groups <- 5
num_per_group <- 20
```

## Generate or Load Random Graph

```{r random-graph}
random_graph_file <- paste0(
  "../GRACKLE_results/small_simulation_A/random_graph_", num_nodes, "nodes.RData"
)

if (file.exists(random_graph_file)) {
  message("Loading random graph")
  load(random_graph_file)
} else {
  message("Generating random graph")
  random_graph <- old_randomNetwork(prior_graph = breast_graph, num_nodes = num_nodes)
  save(random_graph, file = random_graph_file)
}
```

## Adjust Transitivity and Cluster Network

```{r adjust-transitivity}
for (transitivity in transitivity_sequence) {
  message(paste("Transitivity sequence:", transitivity))
  
  if (transitivity == 0) {
    adjusted_graph <- random_graph
  } else {
    adjusted_graph <- adjustTransitivity(random_graph, target_transitivity = transitivity)
  }
  
  network_transitivity <- transitivity(as.undirected(adjusted_graph), weights = NA)
  message(paste("Network transitivity:", network_transitivity))
  
  set.seed(42)
  clusters <- cluster_louvain(as.undirected(adjusted_graph), weights = NA)
  membership <- clusters$membership
  membership_counts <- table(as.factor(membership))
  print(membership_counts)
  
  large_clusters <- names(membership_counts[order(membership_counts, decreasing = TRUE)][1:5])
  network_modularity <- round(modularity(clusters), 4)
  message(paste("Prior network modularity:", network_modularity))
}
```

## Simulate Noisy Expression Data

```{r simulate-expression}
noisy_expression_file <- paste0(
  "../GRACKLE_results/small_simulation_A/data/noisy_simulations_num_nodes_",
  num_nodes, "_num_groups_", num_groups, "_per_group_", num_per_group, "_t_", transitivity, ".RData"
)

if (file.exists(noisy_expression_file)) {
  message("Loading noisy expression")
  load(noisy_expression_file)
} else {
  set.seed(42)
  message("Creating noisy expression")
  noisy_expression <- lapply(noise_sequence, function(noise) {
    simulated_expression <- mclapply(large_clusters, function(cluster) {
      parallelSimulateExpression(
        random_graph, iterations = 3, max_expression = 100,
        num_samples = num_per_group, select_nodes = which(membership == cluster),
        perturbation_constant = 2.8, noise_constant = noise
      )
    }, mc.cores = 8)
    
    simulated_expression <- lapply(simulated_expression, as.data.frame)
    combined_expression <- do.call(rbind, simulated_expression)
    rownames(combined_expression) <- paste0("sample", seq_len(nrow(combined_expression)))
    combined_expression
  })
  
  names(noisy_expression) <- noise_sequence
  metadata <- simulateMetadata(
    group_size = rep(num_per_group, num_groups),
    group_labels = paste0("subgroup", seq_len(num_groups)),
    row_names = paste0("sample", seq_len(nrow(noisy_expression[[1]])))
  )
  
  message("Saving noisy expression")
  save(noisy_expression, metadata, file = noisy_expression_file)
}
```

## Perform GRACKLE Grid Search

```{r grackle-grid-search}
lambda_1_sequence <- seq(0, 1, 0.1)
lambda_2_sequence <- seq(0, 1, 0.1)
grid_search <- expand.grid(lambda_1 = lambda_1_sequence, lambda_2 = lambda_2_sequence)
grid_search$score <- 0

message("GRACKLE grid search ...")
scores <- lapply(seq_len(nrow(grid_search)), function(i) {
  grackle <- GRACKLE(
    Y = noisy_expression[[1]],  # Example using the first noise level
    net_similarity = as.matrix(as_adjacency_matrix(adjusted_graph)),
    patient_similarity = as.matrix(metadata) %*% t(metadata),
    diff_threshold = 1e-4,
    lambda_1 = grid_search$lambda_1[i],
    lambda_2 = grid_search$lambda_2[i],
    k = num_groups,
    verbose = FALSE,
    error_terms = FALSE,
    beta = 0,
    iterations = 100
  )
  
  evaluationWrapper(
    test_expression = noisy_expression[[1]],  # Example using the first noise level
    test_metadata = metadata,
    H_train = grackle$H,
    k = num_groups,
    clusters = clusters,
    aligned_clusters = large_clusters
  )
})

grid_search$score <- unlist(scores)
message(paste("Avg GRACKLE score:", round(mean(grid_search$score), 3)))
message(paste("Top GRACKLE score:", max(grid_search$score)))
```

## Save Results

```{r save-results}
simulation_results_file <- paste0(
  "../GRACKLE_results/small_simulation_A/updated_simulation_results_num_nodes_",
  num_nodes, "_num_groups_", num_groups, "_num_samples_", num_per_group,
  "_t_", transitivity, "_m_", modularity, ".RData"
)
message(paste("Saving", simulation_results_file))
save(grid_search, file = simulation_results_file)
```

## Conclusion

This tutorial demonstrated the steps to simulate gene expression data, adjust network properties, and evaluate the GRACKLE model. Modify the parameters as needed for your specific use case.
